A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-12 ICLP Paper/samplepaper.tex
additions on problem encoding part

\newcommand{\clingo}{\emph{clingo}\xspace}
\newcommand{\clingodl}{\emph{clingo}[DL]\xspace}
\newcommand{\figspace}{\vspace{-1ex}}
\newcommand{\ksverify}[1]{#1}%{\textcolor{red}{#1}}
\makeatletter
\def\fcapsize@figure{\normalfont\normalsize\rmfamily}
\def\fcapstyle@figure{\normalfont\normalsize\itshape}
\makeatother

@@

\section{Multi-shot JSP Solving}\label{sec:problem}
This section describes our successive optimization approach to
JSP solving by means of multi-shot ASP with \clingodl.
We start with specifying the fact format for JSP instances,
then detail problem decomposition based on earliest starting times of operations,
present our ASP encoding with DL constraints for optimizing the makespan of partial schedules, and
finally outline the incorporation of time window overlapping and compression
into the iterative scheduling process.   
% The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard. In the JSP, there are a set of jobs to be processed on a set of machines. The number of jobs is $n$, and the number of machines is $m$. Each job $J_i$ contains a chain $ O_{i,1}, O_{i,2},...,O_{i,m} $ of operations must be executed on an known order. The routings of the operations are deterministic and known a priori, as are the processing times of each operation on each machine. We consider here the problem of minimizing the total completion time (makespan). The task of the scheduling is to determine the starting time for each operation while optimizing the makespan. The basic assumptions are as follows:
%
% \begin{enumerate}
% 	\item The processing time of each operation is fixed.
% 	\item Machine breakdown preemption of the operations are not allowed.
% 	\item Each machine can process only one operation at a time.
% 	\item Each operation can be executed by only one machine.
% 	\item A machine cannot process more than one operation at a time.
% 	\item The jobs and the machines are available at time $0$.
% \end{enumerate}

\subsection{Problem Instance}\label{subsec:instance}
% ASP has been applied to solve different scheduling problems. For example, ASP is used to solve scheduling problems in the healthcare systems \cite{dodaro2017nurse,dodaro2019asp}. One of these studies aimed to assign nurses to shifts according to some constraints\cite{dodaro2017nurse}. In addition, ASP has been proposed to solve a train scheduling problem while minimizing the train delay\cite{abels2019train}. One of the limitations of ASP mentioned in these studies is the grounding issues while solving a large number of instances. One of the recent works has presented a multi-shot solving for scheduling problem \cite{francescutto2021solving}. This paper aims to use different decomposition methods to solve the JSP using ASP with Difference Logic. 
% 
% \paragraph{Problem Instances.} 
%
Each job in a JSP instance is a sequence of operations with associated
machines and processing times.
Corresponding facts for an example instance with three jobs and three
machines are displayed in Listing~\ref{prg:facts}.
An atom of the form
\lstinline{operation(}$j$\lstinline{,}$s$\lstinline{,}$m$\lstinline{,}$p$\lstinline{)}
denotes that the step~$s$ of job~$j$ needs to be processed by machine~$m$ for $p$ time units.
For example, the second operation of job~\lstinline{3} has a processing time of
\lstinline{3} time units on machine~\lstinline{1},
as specified by the fact
\lstinline{operation(3,}\linebreak[1]\lstinline{2,}\linebreak[1]\lstinline{1,3).}
The operation cannot be performed before the first operation of job~\lstinline{3}
is completed,
and its execution must not intersect with the first operation of job~\lstinline{1}
or the second operation of job~\lstinline{2},
which need to be processed by machine~\lstinline{1} as well.
That is, a schedule for the example instance must determine a sequence in which
to process the three mentioned operations on machine~\lstinline{1},
and likewise for operations sharing machine~\lstinline{2} or~\lstinline{3}, respectively.
%
% In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances. For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}. The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}. For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end}. For instance, \emph{operation(1,1)} is processed by machine 2. 
\lstinputlisting[float=b,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}

@@

\subsection{Problem Decomposition}\label{subsec:decomposition}
%
Since JSP instances are highly combinatorial and the ground representation size
can also become problematic for large real-world scheduling problems, achieving
scalability of complete optimization methods necessitates problem decomposition.
To this end, we consider strategies for partitioning the operations of JSP instances
into balanced time windows \hl{(the number of operations assigned to each time window is equal)}, which respect the operation precedence within jobs and
enable a successive extension of good-quality partial schedules to a global solution.
In the following, we first detail problem decomposition based on the earliest starting times of operations, and then outline further strategies that can be encoded by stratified ASP programs as well.

@@

The last rule in line~\ref{prg:assignment_process:assign:begin} inspects the total
order of operations to partition them into time windows of the size~\lstinline{W}
determined by \lstinline{width(W)},
where only the last time window may possibly include fewer operations in case the
split is uneven.
As the ASP program encoding problem decomposition is stratified, its ground
instantiation can be simplified to (derived) facts, as shown in Listing~\ref{prg:tw}
for our example instance.
Time window numbers from~\lstinline{1} to~\lstinline{n = 2} are given by
the third argument of atoms over \lstinline{window/3}, so that the second operation
of job~\lstinline{3} and the third operation of each job form the time window~\lstinline{2},
while time window~\lstinline{1} consists of the five other operations.
%
\lstinputlisting[float=t,label=prg:tw,caption={Example time windows},numbers=none,xleftmargin=3pt]{listing/TW.lp}

@@

While the rules up to line~\ref{prg:base:order:end} yield atoms of the form
\lstinline{order(}$o_1$\lstinline{,}$o_2$\lstinline{,}$p_1$\lstinline{,}$w$\lstinline{)},
expressing the hard requirement or choice to perform an operation~$o_1$
with processing time~$p_1$ before the operation~$o_2$ belonging to time window~$w$,
the remaining rules assert corresponding DL constraints. % on the starting time of~$o_2$.
To begin with,
the starting times of operations~$o$ from the previous time window $w-1$ (if any)
are in the lines~\ref{prg:diff_log:freez1:begin} and~\ref{prg:diff_log:freez2:begin}
fixed to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
as supplied by reified facts \lstinline{start(}$o$\lstinline{,}$t$\lstinline{,}$w-1$\lstinline{).}
In line~\ref{prg:diff_log:non_nega:begin},
the lower bound~$0$ is asserted for the starting time of the first operation of some job,
included in the time window~$w$ for which a partial schedule is to be determined next.
In addition, constraints reflecting the order of performing operations are imposed 
in line~\ref{prg:diff_log:oper_const:begin},
which concerns operations sharing a machine as well as successor operations within jobs.
Since such constraints trace the sequence of operations in a job, they establish the
earliest starting time, considered for problem decomposition in Section~\ref{subsec:decomposition},
as lower bound for scheduling an operation, and the execution order on the machine
processing the operation can increase its starting time further.
The last rule of the \lstinline{step(w)} subprogram in lines
\ref{prg:diff_log:oper_limit:begin}-\ref{prg:diff_log:oper_limit:end} asserts
that the value for the DL variable \lstinline{makespan} cannot be less than the
completion time of any operation of the time window~$w$.
That is, the least feasible \lstinline{makespan} value provides the
scheduling horizon of a partial schedule for operations of time windows up to~$w$.

\begin{figure}[tb]
  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  %  \vspace{0.1\baselineskip}
  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
    \item % \begin{tabbing}
          If $w= 1$, % \= 
          then $P:=\{\text{\lstinline{base}}\}$;
          % \\
          % \>

          % \hfill
          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                      \mid
                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
          % \end{tabbing}
    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
    \item While there is some answer set $X$ and the time limit per time window is not reached:
      \begin{enumerate}[label=\textbf{\alph*.}]
      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
      \end{enumerate}
    \end{enumerate}
  \end{tcolorbox}
  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
  \end{figure}
  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
%
The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false for making sure that
some (partial) schedule is feasible.
Once some schedule with a horizon $h+1$ is found,
the subprogram \lstinline{optimize(}$h$\lstinline{)} is instantiated on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The process of successively reducing the horizon~$h$ in order to find better partial schedules
is repeated until the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false and the successive optimization proceeds by instantiating the
\lstinline{step(}$w+1$\lstinline{)} subprogram for the next time window $w+1$ (if any), and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution is implemented by means of the Python interface of \clingodl,
where a time limit can be configured to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.
\hl{Figure~{\ref{fig:enum}} summarizes the main steps of the Multi-shot solving controlled by a python script.}

@@

\section{Experiments}\label{sec:experiments}
% This section will show the computational experiments that we conducted to test the performance of our proposed model.
In order to evaluate our multi-shot ASP modulo DL approach to JSP solving,
we ran experiments on JSP benchmark sets by \cite{taillard1993benchmarks}
of several sizes,
considering ten instances each with
$50\times15$, $50\times20$, or $100\times20$ jobs and machines. \hl{We have selected the largest set of instances generated and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
The instances are generated such that each job includes one operation per machine, e.g.,
$15$ operations in case of the $50\times15$ instances,
where the sequence in which the operations of a job are allocated to machines varies.
In our experiments, we assess decomposition strategies presented in Section~\ref{subsec:decomposition}
with different number and size of time windows as well as the impact of
time window overlapping and compression techniques described in Section~\ref{subsec:encoding}.
For the comparability of results between runs with a different number of time windows
and respective optimization subproblems addressed by multi-shot solving,
we divide the total runtime limit of $1000$ seconds for \clingodl (version 1.3.0) by the number
of time windows to evenly spend optimization efforts on subproblems.
Our experiments have been run on an 
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.%
\footnote{The benchmarks and our implementation are available at: \myurl}

>>>>

\newcommand{\clingo}{\emph{clingo}\xspace}
\newcommand{\clingodl}{\emph{clingo}[DL]\xspace}
\newcommand{\figspace}{}%{\vspace{-1ex}}
\newcommand{\ksverify}[1]{#1}%{\textcolor{red}{#1}}
\makeatletter
\def\fcapsize@figure{\normalfont\normalsize\rmfamily}
\def\fcapstyle@figure{\normalfont\normalsize\itshape}
\makeatother

@@

\section{Multi-shot JSP Solving}\label{sec:problem}
This section describes our successive optimization approach to
JSP solving by means of multi-shot ASP with \clingodl.
We start with specifying the fact format for JSP instances,
then detail problem decomposition based on earliest starting times of operations,
present our ASP encoding with DL constraints for optimizing the makespan of partial schedules, and
finally outline the iterative scheduling process 
along with the incorporation of time window overlapping and compression techniques.   
% The Job-shop Scheduling Problem (JSP) has been a complex and combinatorial optimization problem since the 1950s, and it was shown to be NP-hard. In the JSP, there are a set of jobs to be processed on a set of machines. The number of jobs is $n$, and the number of machines is $m$. Each job $J_i$ contains a chain $ O_{i,1}, O_{i,2},...,O_{i,m} $ of operations must be executed on an known order. The routings of the operations are deterministic and known a priori, as are the processing times of each operation on each machine. We consider here the problem of minimizing the total completion time (makespan). The task of the scheduling is to determine the starting time for each operation while optimizing the makespan. The basic assumptions are as follows:
%
% \begin{enumerate}
% 	\item The processing time of each operation is fixed.
% 	\item Machine breakdown preemption of the operations are not allowed.
% 	\item Each machine can process only one operation at a time.
% 	\item Each operation can be executed by only one machine.
% 	\item A machine cannot process more than one operation at a time.
% 	\item The jobs and the machines are available at time $0$.
% \end{enumerate}

\subsection{Problem Instance}\label{subsec:instance}
% ASP has been applied to solve different scheduling problems. For example, ASP is used to solve scheduling problems in the healthcare systems \cite{dodaro2017nurse,dodaro2019asp}. One of these studies aimed to assign nurses to shifts according to some constraints\cite{dodaro2017nurse}. In addition, ASP has been proposed to solve a train scheduling problem while minimizing the train delay\cite{abels2019train}. One of the limitations of ASP mentioned in these studies is the grounding issues while solving a large number of instances. One of the recent works has presented a multi-shot solving for scheduling problem \cite{francescutto2021solving}. This paper aims to use different decomposition methods to solve the JSP using ASP with Difference Logic. 
% 
% \paragraph{Problem Instances.} 
%
Each job in a JSP instance is a sequence of operations with associated
machines and processing times.
Corresponding facts for an example instance with three jobs and three
machines are displayed in Listing~\ref{prg:facts}.
%
% In order to encode the problem using ASP, we should define a set of predicates that represents the problem instances. For instance, the operations are encoded with the predicate \lstinline{operation/2} where the first variable denotes the job number and the second is the operation number, see lines \ref{prg:facts:ops:begin}-\ref{prg:facts:ops:end} in Listing \ref{prg:facts}. The time needed to finish an operation is represented by the predicate \lstinline{pro/3} in which the operation is determined by the first two terms and the third represents the processing time in lines \ref{prg:facts:pro:begin}-\ref{prg:facts:pro:end}. For the machine assignment, the predicate \lstinline{assign/3} shows that a particular operation is executed by a machine in lines \ref{prg:facts:mach:begin}-\ref{prg:facts:mach:end}. For instance, \emph{operation(1,1)} is processed by machine 2. 
\lstinputlisting[float=b,label=prg:facts,caption={Example JSP instance},linerange={1-3},numbers=none,xleftmargin=3pt]{listing/facts.lp}
%
An atom of the form
\lstinline{operation(}$j$\lstinline{,}$s$\lstinline{,}$m$\lstinline{,}$p$\lstinline{)}
denotes that the step~$s$ of job~$j$ needs to be processed by machine~$m$ for $p$ time units.
For example, the second operation of job~\lstinline{3} has a processing time of
\lstinline{3} time units on machine~\lstinline{1},
as specified by the fact
\lstinline{operation(3,}\linebreak[1]\lstinline{2,}\linebreak[1]\lstinline{1,3).}
The operation cannot be performed before the first operation of job~\lstinline{3}
is completed,
and its execution must not intersect with the first operation of job~\lstinline{1}
or the second operation of job~\lstinline{2},
which need to be processed by machine~\lstinline{1} as well.
That is, a schedule for the example instance must determine a sequence in which
to process the three mentioned operations on machine~\lstinline{1},
and likewise for operations sharing machine~\lstinline{2} or~\lstinline{3}, respectively.

@@

\subsection{Problem Decomposition}\label{subsec:decomposition}
%
Since JSP instances are highly combinatorial and the ground representation size
can also become problematic for large real-world scheduling problems, achieving
scalability of complete optimization methods necessitates problem decomposition.
% To this end, 
In order to enable a successive extension of good-quality partial schedules to a global solution,
we consider strategies for partitioning the operations of JSP instances
into balanced time windows, \hl{each comprising an equal number of operations such that their precedence
within jobs is respected}. % , which respect the operation precedence within jobs and
% in order to enable a successive extension of good-quality partial schedules to a global solution.
In the following, we first detail problem decomposition based on the earliest starting times of operations, and then outline further strategies that can be encoded by stratified ASP programs as well.

@@

The last rule in line~\ref{prg:assignment_process:assign:begin} inspects the total
order of operations to partition them into time windows of the size~\lstinline{W}
determined by \lstinline{width(W)},
where only the last time window may possibly include fewer operations in case the
split is uneven.
As the ASP program encoding problem decomposition is stratified, its ground
instantiation can be simplified to (derived) facts, as shown in Listing~\ref{prg:tw}
for our example instance.
Time window numbers from~\lstinline{1} to~\lstinline{n = 2} are given by
the third argument of atoms over \lstinline{window/3}, so that the second operation
of job~\lstinline{3} and the third operation of each job form the time window~\lstinline{2},
while time window~\lstinline{1} consists of the five other operations.
%
\lstinputlisting[float=b,label=prg:tw,caption={Example time windows},numbers=none,xleftmargin=3pt]{listing/TW.lp}

@@

While the rules up to line~\ref{prg:base:order:end} yield atoms of the form
\lstinline{order(}$o_1$\lstinline{,}$o_2$\lstinline{,}$p_1$\lstinline{,}$w$\lstinline{)},
expressing the hard requirement or choice to perform an operation~$o_1$
with processing time~$p_1$ before the operation~$o_2$ belonging to time window~$w$,
the remaining rules assert corresponding DL constraints. % on the starting time of~$o_2$.
To begin with,
the starting times of operations~$o$ from the previous time window $w-1$ (if any)
are in the lines~\ref{prg:diff_log:freez1:begin} and~\ref{prg:diff_log:freez2:begin}
fixed \hl{by restricting them from above or below, respectively,}
to the value~$t$ in an (optimized) partial schedule for time window $w-1$, 
as supplied by reified facts \lstinline{start(}$o$\lstinline{,}$t$\lstinline{,}$w-1$\lstinline{).}
In line~\ref{prg:diff_log:non_nega:begin},
the lower bound~$0$ is asserted for the starting time of the first operation of some job,
included in the time window~$w$ for which a partial schedule is to be determined next.
In addition, constraints reflecting the order of performing operations are imposed 
in line~\ref{prg:diff_log:oper_const:begin},
which concerns operations sharing a machine as well as successor operations within jobs.
Since such constraints trace the sequence of operations in a job, they establish the
earliest starting time, considered for problem decomposition in Section~\ref{subsec:decomposition},
as lower bound for scheduling an operation, and the execution order on the machine
processing the operation can increase its starting time further.
The last rule of the \lstinline{step(w)} subprogram in lines
\ref{prg:diff_log:oper_limit:begin}-\ref{prg:diff_log:oper_limit:end} asserts
that the value for the DL variable \lstinline{makespan} cannot be less than the
completion time of any operation of the time window~$w$.
That is, the least feasible \lstinline{makespan} value provides the
scheduling horizon of a partial schedule for operations of time windows up to~$w$.

The task of optimizing the horizon of a (partial) schedule means choosing
the execution order of operations of the latest time window sharing a machine
such that the value for the \lstinline{makespan} variable is minimized.
In single-shot ASP modulo DL solving with \clingodl, this can be accomplished
via the command-line option \lstinline{--minimize-variable=makespan}.
\hl{This option, however, is implemented by means of a fixed control loop that
cannot be combined with (other) multi-shot solving methods.}
For the successive optimization by time windows, where the scheduling
horizon gradually increases, we thus require a dedicated treatment of DL constraints
limiting the value for \lstinline{makespan}.
To this end, the \lstinline{optimize(m)} subprogram below line~\ref{prg:diff_log:oper_min}
declares an external atom \lstinline{horizon(m)} for controlling whether a DL
constraint asserted in line~\ref{prg:diff_log:oper_min:end} is active and limits
the \lstinline{makespan} value to an integer supplied for the parameter~\lstinline{m}.

\subsection{\hl{Iterative Scheduling}}\label{subsec:scheduling}

\hl{The main steps of our control loop for successive schedule optimization
by multi-shot ASP modulo DL solving, implemented by means of the Python interface of \clingodl,
are displayed in Figure~{\ref{fig:enum}} and further detailed in the following.}
When launching the optimization process for a new time window, any instance of the
\lstinline{horizon(m)} atom introduced before is set to false in step~1 for making sure that
some (partial) schedule is feasible.
\hl{No such atoms have been introduced yet for the first time window $w=1$,
where the static (default) subprogram called} \lstinline{base}\hl{,
supplying a JSP instance along with its decomposition in terms of facts over} \lstinline{window/3}\hl{,
and the} \lstinline{step(}$1$\lstinline{)} \hl{subprogram for operations of the first time window
are instantiated in steps~2 and~3.}
%
\begin{figure}[tb]
  \begin{tcolorbox}[breakable,arc=0mm,outer arc=0mm,colbacktitle=lightgray!50!white,coltitle=black,colframe=gray,fonttitle=\bfseries,colback=white,subtitle style={colback=white}]
  %  \vspace{0.1\baselineskip}
  Starting with $H:=\emptyset$, for $w:= 1$ to number $n$ of time windows:
  \begin{enumerate}[label=\textbf{\arabic*.}]
    \item For each $h\in H$, set \lstinline{horizon(}$h$\lstinline{)} to false
    \item % \begin{tabbing}
          If $w= 1$, % \= 
          then $P:=\{\text{\lstinline{base}}\}$;
          % \\
          % \>

          % \hfill
          else $P := \{\text{\lstinline{start((}$j$\lstinline{,}$s$\lstinline{),}$t$\lstinline{,}$w-1$\lstinline{).}}
                      \mid
                      \{\text{\lstinline{window(}$j$\lstinline{,}$s$\lstinline{,}$w-1$\lstinline{)}},
                      \text{\lstinline{(}$j$\lstinline{,}$s$\lstinline{)}}=t\}\subseteq X\}$
          % \end{tabbing}
    \item Ground $P\cup\{\text{\lstinline{step(}$w$\lstinline{)}}\}$
    \item While there is some answer set $X$ and the time limit per time window is not reached:
      \begin{enumerate}[label=\textbf{\alph*.}]
      \item $h:=t-1$, where $($\lstinline{makespan}${}=t)\in X$
      \item If $h\notin H$, then $H := H\cup\{h\}$ and ground $\{$\lstinline{optimize(}$h$\lstinline{)}$\}$
      \item Set \lstinline{horizon(}$h$\lstinline{)} to true
      \end{enumerate}
    \end{enumerate}
  \end{tcolorbox}
  \caption{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
  \end{figure}%
  % \captionof{figure}{Control loop for successive schedule optimization by multi-shot ASP modulo DL solving\label{fig:enum}}
%
Once some schedule with a horizon $h+1$ is found in step~4a,
the step~4b consists of instantiating
the subprogram \lstinline{optimize(}$h$\lstinline{)} % is instantiated
on demand, i.e.,
in case $h$ has not been passed as a value for~\lstinline{m} before, and
the corresponding \lstinline{horizon(}$h$\lstinline{)} atom is set to true in step~4c for activating the DL
constraint reducing the admitted scheduling horizon to~$h$.
The step 4 of successively reducing the horizon~$h$ in order to find better partial schedules
% is repeated until 
stops when the imposed \lstinline{makespan} value turns out to be infeasible,
meaning that an optimal partial schedule has been found.
As already mentioned above, the introduced instances of the external \lstinline{horizon(m)} atom
are then set to false in step~1, and the successive optimization proceeds by in steps~2 and~3 instantiating the
\lstinline{step(}$w+\nolinebreak1$\lstinline{)} subprogram for the next time window $w+1$ (if any) and
also supplying the determined starting times of operations from time window~$w$
by reified facts.
The described control loop for successively extending good-quality partial schedules to
a global solution 
% is implemented by means of the Python interface of \clingodl,
% where a time limit 
can be configured with a time limit,
included as secondary stopping condition in step~4,
to restrict the optimization efforts per time window
and thus make sure that the iterative scheduling progresses.

@@

\section{Experiments}\label{sec:experiments}
% This section will show the computational experiments that we conducted to test the performance of our proposed model.
In order to evaluate our multi-shot ASP modulo DL approach to JSP solving,
we ran experiments on JSP benchmark sets by \cite{taillard1993benchmarks}
of several sizes,
considering ten instances each with
$50\times15$, $50\times20$, or $100\times20$ jobs and machines. \hl{We have selected the largest set of instances generated and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
The instances are generated such that each job includes one operation per machine, e.g.,
$15$ operations in case of the $50\times15$ instances,
where the sequence in which the operations of a job are allocated to machines varies.
In our experiments, we assess decomposition strategies presented in Section~\ref{subsec:decomposition}
with different number and size of time windows as well as the impact of
time window overlapping and compression techniques described in Section~\ref{subsec:scheduling}.
For the comparability of results between runs with a different number of time windows
and respective optimization subproblems addressed by multi-shot solving,
we divide the total runtime limit of $1000$ seconds for \clingodl (version 1.3.0) by the number
of time windows to evenly spend optimization efforts on subproblems.
Our experiments have been run on an 
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.%
\footnote{The benchmarks and our implementation are available at: \myurl}
